---
title: "intro to viz"
author: "Charles Lang"
date: "October 3, 2016"
output: html_document
---
#Input
```{r}
D1 <- read.csv("School_Demographics_and_Accountability_Snapshot_2006-2012.csv", header = TRUE, sep = ",")

D2 <- subset(D1, schoolyear == 20112012)
```

#Histograms
```{r}
# hist()

hist(D2$frl_percent, breaks = 100)

hist(D2$frl_percent, breaks = c(0,10,20,80,100))

hist(D2$frl_percent, breaks = 100, ylim = c(0,30))
```

#Plots
```{r}
plot(D2$ell_num, D2$ctt_num)

plot(D2$ell_num ~ D2$ctt_num)
# ?plot
#Barplot
x <- c(1,3,2,7,6,4,4)
y <- c(2,4,2,3,2,4,3)

table1 <- table(x,y)

barplot(table1)

#Lineplot
D3 <- aggregate(D1, by = list(D1$schoolyear), FUN = mean)

plot(D3$schoolyear, D3$total_enrollment, type = "l", lty = "dashed")

#Boxplot
D5 <- subset(D1, DBN == "31R075")
D5 <- droplevels(D5)
boxplot(D5$total_enrollment ~ D5$Name)
```
#Pairs
```{r}
D4 <- D2[,c(5,6, 21:24)]

pairs(D4)
```
# Exercise

1. Create a simulated data set containing 100 students, each with a score from 1-100 representing performance in an educational game. The scores should tend to cluster around 75. Also, each student should be given a classification that reflects one of four interest groups: sport, music, nature, literature.
```{r}
stu_num <- c(1:100)
set.seed(123)
score <- rnorm(n=100, mean = 75, sd = 10)
score <- as.integer(score)
score
interest <- sample(c("sport", "music", "nature", "literature"), 100, replace = TRUE)
interest
my_df <- data.frame(stu_num, score, interest)
my_df
```



2. Using base R commands, draw a histogram of the scores. Change the breaks in your histogram until you think they best represent your data.
```{r}
hist(my_df$score, breaks = 100)
```


3. Create a new variable that groups the scores according to the breaks in your histogram.
```{r}
my_df$group <- ifelse(my_df$score >= 90, 1, ifelse(my_df$score >= 80, 2, ifelse(my_df$score >= 70, 3, ifelse(my_df$score >= 60, 4, 5))))
my_df
```


4. Now using the colorbrewer package (RColorBrewer; http://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3) design a pallette and assign it to the groups in your data on the histogram.
The following link is an intro and example of this package
https://moderndata.plot.ly/create-colorful-graphs-in-r-with-rcolorbrewer-and-plotly/
```{r}
library(RColorBrewer)
display.brewer.all()
# Blues is great for me
hist(my_df$group, breaks = c(0:5), col = brewer.pal(5, "Blues"))
# ?hist
# ?boxplot
```


5. Create a boxplot that visualizes the scores for each interest group and color each interest group a different color.
```{r}
# ?boxplot
# boxplot(rand.data,col=brewer.pal(8,"Set3"))
boxplot(score ~ interest, data = my_df, col = brewer.pal(4,"Set1"))
```

6. Now simulate a new variable that describes the number of logins that students made to the educational game. They should vary from 1-25.
```{r}
?sample
set.seed(1)
login = sample(c(1:25), 100, replace = TRUE)
login
my_df <- cbind(my_df, login)
my_df
```

7. Plot the relationships between logins and scores. Give the plot a title and color the dots according to interest group.
```{r}
?plot
# plot(D2$ell_num, D2$ctt_num)
# plot(D2$ell_num ~ D2$ctt_num)
plot(my_df$login, my_df$score, col = factor(my_df$interest), main = "login~score", pch = 16)

```

8. R contains several inbuilt data sets, one of these in called AirPassengers. Plot a line graph of the the airline passengers over time using this data set.
```{r}
# plot(D3$schoolyear, D3$total_enrollment, type = "l", lty = "dashed")
AirPassengers
class(AirPassengers)
plot(AirPassengers, type = "l")
```


9. Using another inbuilt data set, Iris, plot the relationships between all of the variables in the data set. Which of these relationships is it appropraiet to run a correlation on? 
```{r}
pairs(iris)
```

